%\VignetteEngine{knitr::knitr}

# Analyze exit poll for 2016 general presidential election

**Questions**:
  * What characterizes Trump supporters?
  * What motivates Trump supporters?

```{r}
library(io);
library(printr);

cconds <- qread("ctable-prop_class-conditionals.rds");
fconds <- qread("ctable-prop_feature-conditionals.rds");
counts <- qread("ctable-counts.rds");
```

## Are Trump supporters predominately white?
```{r}
cconds[["race"]]
```
86% of Trump supporters are white. According to the 2010 census, the general 
US population consists of 72.4% white Americans.
(http://www.census.gov/prod/cen2010/briefs/c2010br-02.pdf)

## Are Trump supporters predominately male?
```{r}
cconds[["gender"]]
```
54% of Trump supporteres are men.


## Are women more likely to vote for Clinton?
```{r}
fconds[["gender"]]

counts[["gender"]]
fisher.test(as.matrix(counts[["gender"]][, c("trump", "clinton")]))
```
Female voters are 66% more likely to vote for Trump.


## Are Trump supporters predominately 'uneducated'?
```{r}
cconds[["education"]]
```
45% of Trump supporters have a college or postgraduate degree.

## Are Trump female white supporters predominately 'uneducated'?
```{r}
cconds[["education among whites by sex"]]

trump.white.females.ed <- counts[["education among whites by sex"]][1:2, "trump"];
trump.white.females.ed / sum(trump.white.females.ed)
```
46% of white, female Trump supporters have a college degree.


## Are Trump supporters predominately Republicans?
```{r}
cconds[["party id"]]
cconds[["party by gender"]]
```

## Are Trump supporters predominately poor people?
```{r}
cconds[["income"]]
```

## Is the midddle class more likely to vote for Trump?
```{r}
fconds[["income"]]

middle.class <- rbind(
	middle_class = counts[["income"]][3, c("trump", "clinton")],
	not_middle_class = colSums(counts[["income"]][-3, c("trump", "clinton")])
);

middle.class
fisher.test(middle.class)
```
Those earning between $50,000 and $100,000 per year are 20% more likely to
vote for Trump.


## Are Trump supporters mostly married?
```{r}
cconds[["marital status"]]
```

## Are Trump supporters mostly conservatives?
```{r}
cconds[["ideology"]]
```

## Are Trump supporters predominately religious?
```{r}
cconds[["religion"]]
cconds[["how often do you attend religious services?"]]
```

## Are Trump supporters predominately Evangelical Christians?
```{r}
cconds[["white born-again or evangelical christian?"]]
```

## Do Trump supporters predominately live in rural areas?
```{r}
cconds[["area type"]]
```


## Which issues do Trump supporters care about?
```{r}
cconds[["most important issue facing the country"]]
```

## What do Trump supporters think of international trade?
```{r}
cconds[["effect of international trade"]]
```

## What do Trump supporters think of the ongoing war?
```{r}
cconds[["how is the fight against isis going?"]]
```

## What do Trump supporters think of Obamacare?
```{r}
cconds[["view on obamacare"]]
```

## Do Trump supporters want the wall?
```{r}
cconds[["view of u.s. wall along the entire mexican border"]]
```
79% of Trump supporters support the construction of the wall.

## How has the financial situation of Trump supporters changed compared to four years ago?
```{r}
cconds[["financial situation compared to four years ago:"]]
```
44% of Trump supporters are financially worse today compared to four years
ago.


## Which candidate qualities matter the most to Trump supporters?
```{r}
cconds[["which candidate quality mattered most?"]]
```
73% of Trump supporters want change.

## What do Trump supporters think of Obama?
```{r}
cconds[["opinion of barack obama as president"]]
```

## What do Trump supporters think of Hillary?
```{r}
cconds[["opinion of hillary clinton"]]
```

## What do Trump supporters think of Trump?
```{r}
cconds[["opinion of donald trump"]]
```

## Does Trump's treatment of women bother Trump supporters?
```{r}
cconds[["does donald trump's treatment of women bother you:"]]
```

## Do Trump supporters believe Trump is qualified to be president?
```{r}
cconds[["is trump qualified to serve as president?"]]
```
23% of Trump supporters do not believe Trump to be qualified to be president.

## Do Trump supporters believe Trump has the temperament to be president?
```{r}
cconds[["does trump have the temperament to be president?"]]
```
28% of Trump suppporters do not believe Trump has the temperatment to be
president.

## Which candidates do Trump supporter believe has the qualification or temperament to be president?
```{r}
cconds[["who is qualified to serve as president?"]]
cconds[["which candidate has the right temperament?"]]
```
21% of Trump supporters believe neither candidate is qualified.
22% of Trump supporters believe neither candidate has the right temperament.

```{r}
cconds[["which candidate is honest?"]]

honest <- rowSums(counts[["which candidate is honest?"]], na.rm=TRUE);
honest["neither is"] / sum(honest)
```
About 30% of voters believe neither candidate is honest.


```{r}
cconds[["in your vote, were supreme court appointments:"]]

cconds[["does the country's criminal justice system:"]]

cconds[["feelings about the federal government"]]

cconds[["opinion of government"]]

cconds[["is hillary clinton honest and trustworthy?"]]

cconds[["is donald trump honest and trustworthy?"]]

cconds[["is clinton qualified to serve as president?"]]

cconds[["who would better handle the economy?"]]

cconds[["who would better handle foreign policy?"]]

cconds[["who would be a better commander in chief?"]]

cconds[["opinion of the democratic party"]]

cconds[["opinion of the republican party"]]

cconds[["condition of national economy"]]


cconds[["direction of the country"]]

cconds[["life for the next generation of americans will be:"]]

cconds[["importance of debates to your vote"]]

cconds[["were debates a factor in your vote?"]]
```

